---
title: "NYPD Incidents"
output:
  html_document: default
  pdf_document: default 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Packages and libraries

First of all, all the librarys and requires are placed for easier detection by the reader


```{r nypd.csv}

library(lubridate)
library(dplyr)
library(tinytex)
require(ggplot2)
library(leaflet)
library(tidyverse)


```
RR
## NYC Shooting Incidents Analysis

### Cleaning Data

 Shooting incidents are registered by police departament with a lot of information. Many of this information are not going to be used in the analysis of the incidents. With the original file, we can apply the "summary()" function to obtain some information about the data.
 
 
 
  |           Victim Race          | Numer of Incidents |
|:------------------------------:|--------------------|
| American Indian/Alaskan Native |                  9 |
| Asian/Pacific Islander         |                320 |
| Black                          |              16846 |
| Black Hispanic                 |               2244 |
| White Hispanic                 |               3432 |
| White                          |                615 |


The summary data shows us that the number of incidents wich black people are victms are 27 times more than white people. 

Rstudio provides some easy to implement packages to clean the dataframe and manipulate it. First, 3 new dataframes had been created. First of them is to implement a map, the second is to plot historical register of incidents and the last to be filtered by bourogh. The third is also used to bar plot incidents by year.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}  
#  Read csv data
nypd <- read.csv("nypd.csv" , header = TRUE) 
nypd2 <- read.csv("nypd.csv" , header = TRUE)
nypd3 <- read.csv("nypd.csv" , header = TRUE)

#  New York File map
nyfmap <- nypd3 %>%
  #REMOVE COLUMNS
  select(-c('INCIDENT_KEY',  Lon_Lat, JURISDICTION_CODE)) 
    #SORT BY OCCUR DATE THEN OCCUR TIME
    nyfmap <- arrange(nyfmap, OCCUR_DATE, OCCUR_TIME) 
    # TRANSFORM DATE FROM <CHAR> TYPE TO <DATE> TYPE
    nyfmap <- nyfmap[order(as.Date(nyfmap$OCCUR_DATE, format = "%m/%d/%Y")),]%>% 
    mutate(date=mdy(OCCUR_DATE))
    


#  New York File Cleaner
nyfclean <- nypd %>%
  #REMOVE COLUMNS
  select(-c('INCIDENT_KEY', X_COORD_CD, Y_COORD_CD, Latitude,
            Longitude, Lon_Lat, JURISDICTION_CODE)) 
    #SORT BY OCCUR DATE THEN OCCUR TIME
    nyfclean <- arrange(nyfclean, OCCUR_DATE, OCCUR_TIME) 
    # TRANSFORM DATE FROM <CHAR> TYPE TO <DATE> TYPE
    nyfclean <- nyfclean[order(as.Date(nyfclean$OCCUR_DATE, format = "%m/%d/%Y")),]%>% 
    mutate(date=mdy(OCCUR_DATE))
```
    
### Transform Data in A Time Series Data

A time series data is categorized by data points indexed in time order. After this process, the data is ready to be ploted.   
  
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}   
#  Sort the incidents by day and quantify the cases. 
#Important to transform the date column from char, to date type.      
nyf_incidents_by_date<- nypd2 %>%
  select(-c(INCIDENT_KEY, PRECINCT, JURISDICTION_CODE, 
            LOCATION_DESC, STATISTICAL_MURDER_FLAG, PERP_AGE_GROUP, 
            PERP_SEX, PERP_RACE, X_COORD_CD, Y_COORD_CD, Latitude,
            Longitude, Lon_Lat))%>%
    mutate(OCCUR_DATE=mdy(OCCUR_DATE))

  #ADD A UNITY COLUMN TO FACILITATE THE 
  #SUM OF SHOOTING INCIDENTS IN EACH DATE OF THE YEAR
  nyf_incidents_by_date$unity <- rep(1,nrow(nyf_incidents_by_date))
  nyf_incidents_by_date <- arrange(nyf_incidents_by_date, OCCUR_DATE, OCCUR_TIME)
  
  #UNIFY THE ROWS WITH THE SAME DATE AND SUM THE UNITY CELL OF THEM,
  #COUNTING THE SHOOTING INCIDENTS EACH DAY
  nyf_incidents_by_date <- group_by(nyf_incidents_by_date, OCCUR_DATE) %>%
    summarize(Cases=sum(unity))
  nyf_incidents_by_date <- nyf_incidents_by_date[order(as.Date(nyf_incidents_by_date$OCCUR_DATE, format = "%m/%d/%Y")),]
  
  
  
  #Sort the incidents with registered deaths by date and make it a time series
nyf_deaths_by_date <- nypd2 %>% filter(STATISTICAL_MURDER_FLAG== "true") %>%
  select(-c(INCIDENT_KEY, PRECINCT, JURISDICTION_CODE, 
            X_COORD_CD, Y_COORD_CD, Latitude,
            Longitude, Lon_Lat))
  nyf_deaths_by_date <- arrange(nyf_deaths_by_date,
                                OCCUR_DATE, OCCUR_TIME)
  nyf_deaths_by_date$unity <- rep(1,nrow(nyf_deaths_by_date))
  nyf_deaths_by_date <- group_by(nyf_deaths_by_date, OCCUR_DATE) %>%
    summarize(Cases=sum(unity))
  nyf_deaths_by_date <- nyf_deaths_by_date[order(as.Date(nyf_deaths_by_date$OCCUR_DATE, format = "%m/%d/%Y")),]
  
```
 
 
## Boroughs shooting incident analysis
 
 The borough analysis depends on changes in the previous used data. Grouping the incidents by borough, and arranging then by date, a time series data of each borough is formed. The same strategy used previously can be used again, using summary function, it is able to see wich people is more affected by the violence.


```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)} 
 #  Sort the incidents by borough  
nyf_by_boro <- nypd %>%
  select(-c('INCIDENT_KEY', X_COORD_CD, Y_COORD_CD, Latitude,
            Longitude, Lon_Lat, JURISDICTION_CODE, LOCATION_DESC, PRECINCT))
  nyf_by_boro <- nyf_by_boro[order(as.Date(nyf_by_boro$OCCUR_DATE, format = "%m/%d/%Y")),]%>%
  mutate(OCCUR_DATE=mdy(OCCUR_DATE))
  nyf_by_boro$unity <- rep(1,nrow(nyf_by_boro))
  nyf_by_boro <- arrange(nyf_by_boro, BORO, OCCUR_DATE)
    
 #Create a new dataframe for each district, sorted by date 
nyf_bronx <- nyf_by_boro %>% filter(BORO == "BRONX")
nyf_brooklyn <- nyf_by_boro %>% filter(BORO == "BROOKLYN")  
nyf_manhattan <- nyf_by_boro%>% filter(BORO == "MANHATTAN")  
nyf_queens <- nyf_by_boro %>% filter(BORO == "QUEENS")
nyf_si <- nyf_by_boro %>% filter(BORO == "STATEN ISLAND")  

#count the incidents by day in each district data frame
nyf_bronx <- group_by(nyf_bronx, OCCUR_DATE) %>%
  summarize(Cases=sum(unity))
nyf_brooklyn <- group_by(nyf_brooklyn, OCCUR_DATE) %>%
  summarize(Cases=sum(unity))
nyf_manhattan <- group_by(nyf_manhattan, OCCUR_DATE) %>%
  summarize(Cases=sum(unity))
nyf_queens <- group_by(nyf_queens, OCCUR_DATE) %>%
  summarize(Cases=sum(unity))
nyf_si <- group_by(nyf_si, OCCUR_DATE) %>%
  summarize(Cases=sum(unity))
```
 
## Plots 

### Shooting Incidents in NYC by day

```{r , echo=FALSE}
# Plot Number of incidents by date 
nyf_incidents_by_date %>%
  ggplot(aes(x=OCCUR_DATE, y=Cases)) +
  geom_line(aes(color = "Cases")) +
  theme_minimal()+
  ggtitle("Shooting incident in NYC")
```
```{r }
nyf_incidents_by_date %>%
  slice_max(Cases)
```

---
title: "Rmarkdown HTML including Leaflet"
output: html_document
---

Show "Leaflet for R" within html: works.

```{r echo=FALSE}
m <- leaflet(nyfmap) %>%
  addTiles() %>%
  setView(lng = -73.968285, lat= 40.785091, zoom=16) %>%
  addMarkers(clusterOptions = markerClusterOptions(), lng = nyfmap$Longitude, lat= nyfmap$Latitude)

m
```

### Shooting Incidents in NYC by Year


```{r , echo=FALSE}
# barplot
nyf_incidents_by_year <- nyf_incidents_by_date  %>%
  mutate(OCCUR_DATE = format(OCCUR_DATE, "20%y")) 
  nyf_incidents_by_year <-group_by(nyf_incidents_by_year, OCCUR_DATE) %>%
    summarize(Cases = sum(Cases))
  
table1<-table(nyf_incidents_by_year$Cases) %>%



  barplot(height = nyf_incidents_by_year$Cases ,table(nyf_incidents_by_year$OCCUR_DATE),
          main="Shooting Incidents by Year",
          names.arg= nyf_incidents_by_year$OCCUR_DATE,
          xlab= "Year",
          ylab="Count",
          border="red",
          col="blue",
          density=50) 
```


### Shooting Incidents in Bronx by day

```{r, echo=FALSE}
# Plot Number of incidents by date 
nyf_brooklyn %>%
  ggplot(aes(x=OCCUR_DATE, y=Cases)) +
  geom_line(aes(color = "Cases"))+
  ggtitle("Shooting incident in Brooklyn")
```
```{r }
nyf_brooklyn %>%
  slice_max(Cases)
```

### Shooting Incidents in Manhattan by day

```{r , echo=FALSE}
# Plot Number of incidents by date 
nyf_manhattan %>%
  ggplot(aes(x=OCCUR_DATE, y=Cases)) +
  geom_line(aes(color = "Cases"))+
  ggtitle("Shooting incident in Manhattan")
```
```{r }
nyf_manhattan %>%
  slice_max(Cases)
```

### Shooting Incidents in Queens by day
```{r , echo=FALSE}
# Plot Number of incidents by date 
nyf_queens %>%
  ggplot(aes(x=OCCUR_DATE, y=Cases)) +
  geom_line(aes(color = "Cases"))+
  ggtitle("Shooting incident in Queens")
```
```{r }
nyf_queens %>%
  slice_max(Cases)
```

### Shooting Incidents in Staten Island by day
```{r , echo=FALSE}
# Plot Number of incidents by date
nyf_si %>%
  ggplot(aes(x=OCCUR_DATE, y=Cases)) +
  geom_line(aes(color = "Cases"))+
  ggtitle("Shooting incident in Staten Island")
```
```{r }
nyf_si %>%
  slice_max(Cases)
```


## Identifying the Bias

  As a black south american person, with all the social and economic inequality, use to see these problems as a justification of all the violence that happens in my country, Brazil. As said, I tryed to correlate the cited problems to the shooting incidents reported.

## Apply a linear model

A model applyed to data analysis can be helpful to make decisions about the future. As example

```{r, echo = FALSE }
#  Sort the incidents by day and quantify the cases. Important to transform the date column from char, to date type.    
nyf_incidents_by_date<- nypd2 %>%
  select(-c(INCIDENT_KEY, PRECINCT, JURISDICTION_CODE, LOCATION_DESC, STATISTICAL_MURDER_FLAG, PERP_AGE_GROUP, PERP_SEX, PERP_RACE, X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat))%>%
    mutate(OCCUR_DATE=mdy(OCCUR_DATE))
  nyf_incidents_by_date$unity <- rep(1,nrow(nyf_incidents_by_date))
  nyf_incidents_by_date <- arrange(nyf_incidents_by_date, OCCUR_DATE, OCCUR_TIME)
  nyf_incidents_by_date <- group_by(nyf_incidents_by_date, OCCUR_DATE) %>%
    summarize(Cases=sum(unity))
  nyf_incidents_by_date <- nyf_incidents_by_date[order(as.Date(nyf_incidents_by_date$OCCUR_DATE, format = "%m/%d/%Y")),]
  
    
#Sort the number of deaths in the incidents registered by nypd.
nyf_deaths <- nypd2 %>% filter(STATISTICAL_MURDER_FLAG== "true")  
 

#Deaths by date
nyf_deaths_by_date <- nypd2 %>% filter(STATISTICAL_MURDER_FLAG== "true") %>%
  select(-c(INCIDENT_KEY, PRECINCT, JURISDICTION_CODE, X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat)) %>%
    mutate(OCCUR_DATE=mdy(OCCUR_DATE))
  nyf_deaths_by_date <- arrange(nyf_deaths_by_date, OCCUR_DATE, OCCUR_TIME)
  nyf_deaths_by_date$unity <- rep(1,nrow(nyf_deaths_by_date))
  nyf_deaths_by_date <- group_by(nyf_deaths_by_date, OCCUR_DATE) %>%
    summarize(Cases=sum(unity))
  nyf_deaths_by_date <- nyf_deaths_by_date[order(as.Date(nyf_deaths_by_date$OCCUR_DATE, format = "%m/%d/%Y")),]

nyf_incidents_by_year <- nyf_incidents_by_date  %>%
  mutate(OCCUR_DATE = format(OCCUR_DATE, "20%y")) 
  nyf_incidents_by_year <-group_by(nyf_incidents_by_year, OCCUR_DATE) %>%
    summarize(Cases = sum(Cases))
  
nyf_deaths_by_year <- nyf_deaths_by_date  %>%
  mutate(Date = format(OCCUR_DATE, "20%y")) 
  nyf_deaths_by_year <- group_by(nyf_deaths_by_year, Date) %>%
    summarize(Cases = sum(Cases))
  
nyfcomp <- nyf_incidents_by_year
  nyfcomp <- cbind(nyfcomp, Deaths = nyf_deaths_by_year$Cases)

nyfcomp

```

```{r, echo=FALSE}
mod1 <- lm(Cases ~ Deaths, data = nyfcomp)
summary(mod1)
```
```{r }
nyfcomp <- nyf_incidents_by_year
  nyfcomp <- cbind(nyfcomp, Deaths = nyf_deaths_by_year$Cases)
  
 nyfcomp1 <- nyfcomp %>%
   mutate(pred=predict(mod1))
 nyfcomp1 %>% ggplot()+
   geom_point(aes(x=Deaths, y=Cases), color = "blue")+
   geom_point(aes(x=Deaths, y=pred), color = "red")
 ```

 
 
  As shown in summary function, the number of incidents are 244 times the number of deaths. Linear modeling seems to be able to be used only looking at anual count of incidents. Looking at the daily register, as plot with each bourough, the data cannot be categorized by any linear function.
 
 Another observation is that incidents wich black people are involved are greater than any other race in United States. That shows a social inequality in education, opportunities and cultural of people geographically close of each other, only based on wealthy and skin collour.
 

   

